---
title: "Temperature and Autothefts"
author: "Shuvam Das, Ziwen Han, Aidan Lau, Ryan Pelchat, STA130 TUT0207, Groupnum=???"
subtitle: The subtitle of my project
output:
  beamer_presentation: default
  ioslides_presentation: default
  widescreen: yes
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
#loading proper libraries
library(tidyverse)
library(knitr)
# loading neighbourhood data
neighbourhood <- read.csv(file="2016neigh.csv") %>% select(Hood_ID, pop_2016)
#loading autothefts data and joining with neighbourhood data
thefts <- read.csv(file="auto_thefts.csv") %>% select(occurrencedate, Neighbourhood, Hood_ID) %>% full_join(y=neighbourhood, key="Hood_ID")
#loading weather data from 5 years
T14 <- read.csv(file="2014 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
T15 <- read.csv(file="2015 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
T16 <- read.csv(file="2016 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
T17 <- read.csv(file="2017 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
T18 <- read.csv(file="2018 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
#joining weather data into one data frame and removing NA values
temperature <- rbind(T14, T15, T16, T17, T18)
temperature <- temperature %>% mutate(Mean.Temp.C = round(Mean.Temp.C))
temperature
#joining weather data (temperature) with thefts by date
```
Data wrangling starts here!
```{r}
date <- rep(NA, 18178)
for (i in 1:18178){
  date[i] <- substring(thefts$occurrencedate[i],1,10)
}
```


```{r}
#cleaing up thefts data more
thefts <- thefts %>% mutate(date=date) %>% rename(Date.Time=date) %>% select(Neighbourhood, Hood_ID,pop_2016,Date.Time)
```


```{r}
#joining thefts to temperature that day
autothefts <- inner_join(x=thefts,y=temperature, key="Date.Time") %>% filter(!is.na(Mean.Temp.C))
```

```{r}
#add a new variable to classify by high/low crime rate neighbourhood (rate per 1000) 
crimerate <- autothefts %>% group_by(Hood_ID) %>% summarise(rate=1000*n()/mean(pop_2016))
autothefts <- full_join(x=autothefts,y=crimerate,key="Hood_ID")
autothefts <- autothefts %>% mutate(crimerate=ifelse(rate>=median(rate),"High","Low"))
```
```{r}
#histogram of crime rate distribution
ggplot(data=autothefts) + aes(x=rate) + geom_histogram(bins=100, color="black", fill="blue")
#right skewed data so we use median
```
```{r}
#histogram of temperature distributions
ggplot(data=autothefts) + aes(x=Mean.Temp.C) + geom_histogram(bins=50, color="black", fill="red")
#left skew so we use median again
```


```{r}
#add a new variable to classify by hot and cold days by median 
autothefts <- autothefts %>% mutate(hot_cold=ifelse(Mean.Temp.C>=median(Mean.Temp.C), "Hot", "Cold"))
```

```{r}
#State H0: there is no difference, we sim under this assumption
#State HA: there is a difference (significant)
#select only high risk inputs
riskycars <- autothefts %>% filter(crimerate=="High")
#calculate test statistic as difference in crime rate on hot and cold days
riskyhotcoldcars <- riskycars %>% group_by(hot_cold) %>% summarise(means=mean(rate))
riskyteststat <- diff(riskyhotcoldcars$means) %>% as.numeric()
#simulate the difference in crime rate(per 1000) on hot vs cold days by shuffling labels
reps1 <- 5000
sim_diff <- rep(NA,reps1)
set.seed(130)
for (i in 1:reps1){
  sim <- riskycars %>% mutate(hot_cold=sample(hot_cold)) %>% group_by(hot_cold) %>% summarise(means=mean(rate)) %>% summarise(diff(means)) %>% as.numeric()
  sim_diff[i] <- sim
}
```


```{r}
ggplot(data_frame(data=sim_diff)) + aes(x=data) + geom_histogram(bins=30, color="black", fill="grey") + geom_vline(xintercept=riskyteststat, color="red") + geom_vline(xintercept=-riskyteststat, color="red")
#Graph distribution and show p value

```
```{r}
#calculate p value
pvaluerisky <- data_frame(data=sim_diff) %>% filter(data<=riskyteststat|data>=-riskyteststat) %>% 
summarise(pvalue=n()/reps1)
pvaluerisky
#make a conclusion
```


```{r}
#do the same thing as above for low risk neighbourhoods
#State H0: there is no difference
#State HA: there is a difference (significant)
#select only low risk inputs
notriskycars <- autothefts %>% filter(crimerate=="Low")
#calculate test statistic as difference in crime rate on hot and cold days
notriskyhotcoldcars <- notriskycars %>% group_by(hot_cold) %>% summarise(means=mean(rate))
notriskyteststat <- diff(notriskyhotcoldcars$means) %>% as.numeric()
#simulate the difference in crime rate(per 1000) on hot vs cold days by shuffling labels
sim_diff2 <- rep(NA,reps1)
set.seed(130)
for (i in 1:reps1){
  sim <- notriskycars %>% mutate(hot_cold=sample(hot_cold)) %>% group_by(hot_cold) %>% summarise(means=mean(rate)) %>% summarise(diff(means)) %>% as.numeric()
  sim_diff2[i] <- sim
}
```


```{r}
ggplot(data_frame(data=sim_diff2)) + aes(x=data) + geom_histogram(bins=30, color="black", fill="grey") + geom_vline(xintercept=notriskyteststat, color="red") + geom_vline(xintercept=-notriskyteststat, color="red")
#Graph distribution and show p value
```



```{r}
#calculate p value
pvaluenotrisky <- data_frame(data=sim_diff2) %>% filter(data<=notriskyteststat|data>=-notriskyteststat) %>% 
summarise(pvalue=n()/reps1)
pvaluenotrisky
#make a conclusion
```

```{r}
#trend with error
#name axis
ggplot(data=autothefts) + aes(x=Mean.Temp.C, y=rate) + geom_smooth() 
#+ scale_y_discrete(breaks = c(0,10,20,30,40,50), labels=c("0","10","20","30","40","50"), limits=c(0,50))#we have to explain this
```


```{r}
#scatterplot
ggplot(data=autothefts) + aes(x=Mean.Temp.C, y=rate, color = Hood_ID) + geom_point()
```

# A Section Heading on its own page

## A heading that starts on a new page

This document shows a few basics on making slides with R markdown.

To produce the slides, Knit to HMTL (ioslides).

(You can also knit to PDF if you have LaTeX installed on your computer.)


---

Three dashes starts a new page when it is not started by a new header.

---

### This is a heading that doesn't start a new page

You can make text **bold** or in *italics*.

To make bullet points, start the points after a blank line:
 
- point one
- another point


---

Include some R output

```{r}
ggplot(cars, aes(x=speed, y=dist)) + geom_point()
```

---

Include the output without showing the code and R messages (which is what you want for your poster presentation).  This R code chunk also changes the size of the plot.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
#ggplot(cars, aes(x=speed, y=dist)) + geom_point()

```


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
pred_probs1 <- predict(object = tree, newdata = test, type = "prob")
 #pred_probs2 <- predict(object = tree2, newdata = test, type = "prob")
# 
# # Create R objects which will be used to create the ROC plot
# ### TREE 1 ###
 pred1 <- ROCR::prediction(predictions = pred_probs1[,2],
                          labels=test$crimerate)
 perf1 <- ROCR::performance(pred1, 'tpr', 'fpr')
 # Create a data framewith tpr and fpr -> need to use data.frame here
 perf_df1 <- data.frame(perf1@x.values, perf1@y.values)
 names(perf_df1) <- c("fpr", "tpr")

# 
# # Data to plot
 plot_dat <- cbind(rbind(perf_df1), 
                   model = c(rep("Tree B",nrow(perf_df1))))
# 
# 
# # Plot the ROC curve
 ggplot(data = plot_dat, aes(x = fpr, y = tpr, colour = model)) + 
   geom_line() + geom_abline(intercept = 0, slope = 1, lty = 3) + 
   ylab(perf1@y.name) + 
   xlab(perf1@x.name)
```



```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}

#summary(lm(rate~Mean.Temp.C,data=train))$coefficients

reg <- lm(rate~ ( Mean.Temp.C) * (hot_cold + crimerate), data=train)
reg

summary(reg)$r.squared

#ggplot(data = autothefts )+ aes(x = hot_cold * crimerate, y = rate) + geom_point()
```

This plot shows that there is a positive relationship between distance and speed.

## Read more

There is more information on R markdown on the course website at [https://ntaback.github.io/UofT_STA130/Rmarkdownforclassreports.html](https://ntaback.github.io/UofT_STA130/Rmarkdownforclassreports.html)

For more on creating an `ioslides` presentation see [https://rmarkdown.rstudio.com/ioslides_presentation_format.html](https://rmarkdown.rstudio.com/ioslides_presentation_format.html)

# Headings you should include in your project

## Introduction

Include here a few sentences to introduce the problem and provide context. You might want to briefly summarize the data in words (what is the data and what is it used for). You can present the questions you are investigating here. 


## Objectives (optional)

You can list the questions of interest in complete English sentences here to highlight them. 

## Data Summary (optional)

Here you can explain how you cleaned the data and created variables suitable for answering your questions. 
You can also include graphical displays that either motivated or address the questions.

## Statistical Methods

Describe here what you have done to the data without presenting any results (output). If you want to indicate variables by symbols or variable names, define them here. 

## Results

Present the main results here, in order of importance, related to the questions asked. You might use tables or graphs, or other ways to summarize your results.

## Conclusion

Give your main conclusions here. Follow the order of questions you presented. 

Here you can also mention any additional considerations, concerns, or issues you might have. For example, if the results were unexpected, you can discuss this and perhaps offer possible explanations.

## Acknowledgements (optional)

If you received any help from someone other than your team members you can acknowledge them. For example:   
*The authors would like to thank "TA name" for helpful suggestions and comments that improved the presentation of this poster.*

```{r}
autothefts <- autothefts %>% mutate(Mean.Temp.C=round(Mean.Temp.C))
autothefts %>% group_by(Mean.Temp.C) %>% summarise(n=n()) %>%
  ggplot(aes(x=Mean.Temp.C,y=n)) + geom_point()
# 
# 
# library(tidyverse)
# library(knitr)
# library(rpart)
# library(partykit)
# opts_chunk$set(warning = FALSE, message = FALSE)
# set.seed(2072)
# n <- nrow(autothefts)
# training_indices<- sample(1:n, size = round(0.8*n))
# train <- autothefts[training_indices,]
# #?rpart
# test <- autothefts[-training_indices,]
# tree <- rpart(crimerate ~ Mean.Temp.C, data = train)
# 
# plot(as.party(tree), type = "simple", gp = gpar(cex=0.8))
# 
# 
# prediction <- predict(object = tree, newdata = test, type = "prob")
# table(prediction[,2] >= 0.8, test$crimerate)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
#loading proper libraries
library(tidyverse)
library(knitr)
# loading neighbourhood data
neighbourhood <- read.csv(file="2016neigh.csv") %>% select(Hood_ID, pop_2016)
#loading autothefts data and joining with neighbourhood data
thefts <- read.csv(file="auto_thefts.csv") %>% select(occurrencedate, Neighbourhood, Hood_ID) %>% full_join(y=neighbourhood, key="Hood_ID")
#loading weather data from 5 years
T14 <- read.csv(file="2014 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
T15 <- read.csv(file="2015 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
T16 <- read.csv(file="2016 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
T17 <- read.csv(file="2017 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
T18 <- read.csv(file="2018 Toronto.csv") %>% select(Date.Time,Mean.Temp.C)
#joining weather data into one data frame and removing NA values
temperature <- rbind(T14, T15, T16, T17, T18)
temp_day <- temperature %>% mutate(Mean.Temp.C = round(Mean.Temp.C))
temp_day <- temp_day %>% filter(!is.na(Mean.Temp.C)) %>% group_by(Mean.Temp.C) %>% summarise(no_of_days = n())

date <- rep(NA, 18178)
for (i in 1:18178){
  date[i] <- substring(thefts$occurrencedate[i],1,10)
}
```


```{r}
#cleaing up thefts data more
thefts <- thefts %>% mutate(date=date) %>% rename(Date.Time=date) %>% select(Neighbourhood, Hood_ID,pop_2016,Date.Time) %>% filter(Hood_ID !=1) %>%filter(Hood_ID !=21)
```


```{r}
#joining thefts to temperature that day
autothefts <- inner_join(x=thefts,y=temperature, key="Date.Time") %>% filter(!is.na(Mean.Temp.C))

autothefts <- autothefts %>% mutate(Mean.Temp.C=round(Mean.Temp.C))
autothefts <- autothefts %>% group_by(Mean.Temp.C) %>% summarise(no_of_crimes=n()) 
#

autothefts <- full_join(x = autothefts, y = temp_day, key ="Mean.Temp.C" )

autothefts <- autothefts %>% mutate(no_of_crime_per_day_temp = no_of_crimes/no_of_days) 

```


```{r}
#autothefts

# reg <- lm(no_of_crime_per_day_temp~ Mean.Temp.C, data=autothefts)
# cor(x = autothefts$Mean.Temp.C, y = autothefts$no_of_crime_per_day_temp)
# summary(reg)$coefficients

ggplot(data=autothefts) +  aes(x=Mean.Temp.C,y=no_of_crime_per_day_temp) + geom_point() + geom_smooth(method = 'lm', se= FALSE)

reg <- lm(no_of_crime_per_day_temp~ Mean.Temp.C, data=autothefts)
summary(reg)$r.squared
summary(reg)$coefficients

#joining weather data (temperature) with thefts by date
```

```{r}
set.seed(2072)
n <- nrow(autothefts)
training_indices<- sample(1:n, size = round(0.5*n))
train <- autothefts[training_indices,]
y_train <- train$no_of_crime_per_day_temp
test <- autothefts[-training_indices,]
y_test <- test$no_of_crime_per_day_temp

model_train <- lm(no_of_crime_per_day_temp ~ Mean.Temp.C, data=train)

yhattrain <- predict(model_train, newdata=train)
yhattest <- predict(model_train, newdata=test)

trainRMSE <- sqrt(sum((y_train-yhattrain)^2)/nrow(train))
testRMSE <- sqrt(sum((y_test-yhattest)^2)/nrow(test))

data_frame(Model=c("Predictions"), RMSEtrain=c(trainRMSE), RMSEtest=c(testRMSE))
```

